\textbf{Advection versus diffusion in brain intraventricular transport}

### 2025-03-27 paper/main.tex
minor udpates

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b, c.} Particle tracking data visualizing CSF flow directionality (\textbf{b}) and
    velocity magnitudes (\textbf{c}) in the middle ventricle.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue). Dorsal (D), ventral (V), anterior (A), posterior (P),
    left (L) and right (R) coordinates indicated.
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-only/no-cardiac flow model.
    \textbf{f.} Streamlines of the CSF flow field simulated with the cardiac-only/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{g.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries.
    \textbf{h--j.} Streamlines simulated with the baseline flow model at
    a time a time 25\% (\textbf{h}), 50\% (\textbf{i}) and 75\% (\textbf{j})
    into the cardiac cycle. Times indicated in panel \textbf{g}.
    %t=0.112613 s %t=0.225225 s %t=0.315315 s
    Scale bars 50 \textmu m.
    }
    \label{fig:fig1}
\end{figure}

@@

Proteins, neurotransmitters and other molecular substances
intrinsically diffuse within the brain ventricles, but are also
advected by the flow of CSF. A natural question is which of these
mechanisms that dominate for any given type of solute. To address this
question, we simulate the distribution and evolution of three
different diffusing solutes injected within a region of interest (ROI 1) in the
dorsal region of the middle ventricle (\Crefpanels{fig:fig2}{a}),
subject to the CSF flow induced by motile cilia and cardiac pressure
pulsations. In particular, we consider three different molecular
diffusion coefficients, corresponding to a large extracellular vesicle ($D_1 = 2.17
\times 10^{-12}$ m$^2$/s), a 90 kDa protein STM-GFP ($D_2 = 5.75
\times 10^{-11}$ m$^2$/s), and a 26 kDa photoconvertible protein
Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).

@@

For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have reduced
mean concentrations. To summarize, our results show that the
ventricular geometry and the cross-sectional area of the ventricular ducts matter
for solute transport, with notable differences in transport resulting from
relatively modest geometry modifications. This may explain why
we observed larger differences in the cilia-deficient computational
model results compared to the experimental data in \Cref{fig:fig4},
as the model and the fish have slightly different geometries.
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios thus
    display $F_i/F_{\mathrm{original}}\times 100\%$.}
    \label{fig:fig6}
\end{figure}

@@

\subsection{Numerical approximation and computational model parameters}
We modeled motile cilia as a stress
exerted tangentially on the CSF at the ventricular walls
while employing a slip boundary condition,
motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles, both previously and in this work, in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the governing equations of the flow
is computationally more expensive
than more conventional approaches such as a
Taylor-Hood discretization
($P_k-P_{k-1}$ elements)~\cite{Stenberg1990ErrorProblem}, or lower-order enriched
elements, such as the mini-element
($P_1-P_1$ with a velocity stabilization)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
This yields $\normlinf{\nabla\cdot\uu_h}=0$ to within solver tolerances,
an important property for the stability of the advection-diffusion
equation\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}.
In addition to its favorable mass-conserving property,
the BDM-DG discretization scheme 
we employed for the Stokes equations
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly without any tailored approach,
since the BDM elements control the normal component of the 
velocity~(\Cref{subsec:appendixA1})~\cite{Brezzi1985TwoProblems}.
If instead (for example) Taylor-Hood
elements were to be used, the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties and increases problem complexity.

The ciliated boundaries in our CSF flow model are regionally homogeneous,
whereas in reality, cilia populate the ventricle walls in a
heterogeneous manner~\cite{Olstad2019CiliaryDevelopment}. Heterogeneous
arrangement of cilia, both in location, beating direction and beating frequency,
has been shown to enhance particle clearance in airway cilia arrays
in mice~\cite{Ramirez-SanJuan2020Multi-scaleArrays}.
Incorporating spatial arrangement heterogeneity in our model could possibly
result in different transport timescales,
but we would not expect notable changes to the general trends of the macroscopic
transport that we study here, since the effects of introducing heterogeneous spacing
of cilia would alter the magnitude of cilia forces, but not their direction.
On the other hand, introducing more heterogeneous beating directionality could
alter the transport timescales as well as the directionality.
In our model, all of the applied cilia
tangential traction is aligned with the rostrocaudal axis, and more heterogeneity
in this alignment to reflect heterogeneity in the beating directionality of
cilia would presumably affect both magnitude and directionality of advective transport.
Previous studies have demonstrated that changes to beating directionality affects transport,
and that specific coordination of the ciliary beating is important in the brain ventricles for
flow and signaling processes~\cite{Eichele2020Cilia-drivenVentricle, Olstad2019CiliaryDevelopment,
Faubel2016Cilia-basedVentricles, Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Pellicciotta2020Cilia},
but also in other phenomena such as maintaining healthy functioning of mucous transport in
airways~\cite{Rayner1996CiliarySyndrome, Schneiter2021Multi-scaleFunction,
Bustamante-Marin2019LackClearance, Tsukita2012CoordinatedFeet} and in nodal flow which establishes
left-right symmetry in mammals~\cite{Yoshiba2014RolesSymmetry, Hirokawa2006NodalAsymmetry,
Sawamoto2006NewBrain}.
Although these studies were mostly on spatial scales resolving individual cilia,
we could envision the same effects in our continuum model,
similar to the results of previous numerical studies~\cite{Ramirez-SanJuan2020Multi-scaleArrays,
Thouvenin2020OriginCanal, Yoshida2022EffectVentricles}.

To simulate protein transport in the brain ventricles, we numerically solved
an advection-diffusion equation. Except for where we imposed a normal pressure
boundary condition on the CSF flow,
we imposed a no-flux condition on the concentration. Imposing a pressure-driven
cardiac-related flow on certain parts of the CSF flow boundary leads to in- and outflow,
and the advective transport then requires handling of in- and outflux boundary conditions.
We set an approximate periodic boundary condition on the the in- and outflux boundaries,
such that the total mass is conserved in the
limit that the timestep size is zero, reflecting our assumption that the
ventricular system is closed.
We handled the outflux by imposing zero diffusive flux while
leaving the advective flux term as an unknown in the weak formulation
of the advection-diffusion equation. This has been referred to with different names
in previous works, names such as the ``free boundary condition"~\cite{Papanastasiou1992ACondition},
the ``no boundary condition"~\cite{Griffiths1997TheCondition}, and the 
``consistent flux boundary condition"~\cite{Lynch2020NumericalHemodynamics}. 
In leaving the advective flux term in the weak form we are lacking a boundary condition,
which is problematic in the continuous sense of the problem.
However, the discrete problem is well-posed, because the approach represents
constraining higher-order derivatives of the concentration~\cite{Griffiths1997TheCondition}. 
In defining ROIs 5 and 6 in the anterior and posterior ventricles,
we deliberately excluded the space proximal to
in- and outflow boundaries to limit adverse effects of 
the advective flux boundary condition on the mean concentrations.

A possible improvement to the in- and outflow boundary conditions would be
to consider a 3D-0D coupling of the in- and outflow boundaries using
ordinary differential equations,
similar to what has been done in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
However, if the ventricular system in reality is a closed system,
one should rather seek out a way of modeling the CSF flow without open boundaries to avoid
handling in- and outflow boundary conditions. A more appropriate way of handling the beating
cardiac motion could be to impose a ventricular wall deformation, assumed to result from pulsatile
motion of the heart and/or cardiac vessels surrounding the ventricles. This wall deformation would
induce CSF flow while simultaneously permitting enforcement of the impermeability condition
$\uu\cdot\nn=0$ on the ventricular walls. Such wall motion-induced CSF flow has been modeled and
simulated successfully in the human brain ventricular system~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

The diffusion coefficients used in this study are based on experimental values. 
The value $D_3=\num{1.15e-10} \ \mathrm{m^2/s}$
is based on the value $\num{1.15e-10} \pm 0.11 \ \mathrm{m^2/s}$ reported in an
experiment~\cite{GuraSadovsky2017MeasurementExpansion}, while the Stokes-Einstein
relation~(\Cref{eq:D_stokes_einstein})~\cite{Einstein1905UberTeilchen} was used to extrapolate
the value of $D_2$ and estimate the value of $D_1$, based on experimental
data~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells,
Mahmood2023ExosomeTemperature}. The Stokes-Einstein relation assumes the diffusing
substance is of spherical shape, which proteins are not, and the relation does not
necessarily hold in CSF. Moreover, the extrapolation implies the assumption that
the two substances (the one for which we determine $D$ and the one we extrapolate from)
have equal densities. Consequently, there is uncertainty in the diffusion coefficient values used.
We note that Sadovsky~\emph{et al.}~\cite{GuraSadovsky2017MeasurementExpansion} reported that
Dendra2 satisfies the Stokes-Einstein in a solution of similar viscosity
to CSF. We used CSF density and viscosity values measured in
humans~\cite{Bloomfield1998EffectsFluid},
and because embryonic CSF in zebrafish is high in proteins, nutrients and other substances,
the CSF in which the photoconverted Dendra2 are transported in our experiments may be more viscous
than we assumed. This could reduce the diffusion coefficient (cf.~\Cref{eq:D_stokes_einstein}),
but based on reports of protein concentration having a limited impact on the viscosity of
CSF~\cite{Bloomfield1998EffectsFluid, Brydon1995PhysicalViscosity}, this reduction is presumably
limited. Merely the presence of other proteins and nutrients may also reduce the diffusion
coefficient by reduction of mean-free path lengths of molecules~\cite{Cussler2009Diffusion:Systems}.
Considering the diffusion coefficients $D_1$ and $D_2$ (both smaller
than $D_3$) in our transport analysis provides information about how the simulation
results of Dendra2 transport would be altered if the
diffusion coefficient $D_3$ is overestimated. In the case that $D_3$ is underestimated, 
we expect a trend similar to the difference in the $D_2$ and $D_3$ solute concentrations;
incerasingly diffusive fields.

We heuristically determined the magnitude of the cilia tangential traction boundary condition
with the parameter $\tau = 0.65$ mPa. The maximum stress
$\vert\btau\vert_{\mathrm{max}} = 2.5 \ \mathrm{mPa} = \num{2.5e-3} \ \mathrm{N/m^2}$
was applied in the dorsal-posterior region of the middle ventricle.
Previous simulation studies of cilia-mediated flow in brain ventricles have
used a volumetric force density to represent the cilia.
Based on reconstruction of CSF flow in murine brain ventricles,
Siyahhan~\emph{et al.}~\cite{Siyahhan2014FlowVentricles}
employed a force that increased linearly from the ventricular wall to a maximum force density of
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$ at a distance 15 \textmu m (the cilium length)
away from the wall. If we assumed a cilia array of the same thickness
$\delta = 15$ \textmu m in our model, the maximum stress $\vert\btau\vert_{\mathrm{max}}$
would correspond to a maximum force density of 
\begin{equation*}
    f_c = \frac{\vert\btau\vert_{\mathrm{max}}}{\delta}
        = \frac{\num{2.5e-3} \ \mathrm{N/m^2}}{\num{15e-6} \ \mathrm{m}}
        = 169 \ \mathrm{N/m^3},
\end{equation*}
which is smaller than $f_{\mathrm{max}}$, albeit the same order of magnitude.
The length 15 \textmu m is the typical length of cilia in the human brain
\cite{Afzelius2004CiliaRelatedDiseases, Siyahhan2014FlowVentricles},
while in zebrafish they are closer to 5 \textmu m~\cite{Olstad2019CiliaryDevelopment,
Salman2022ComputationalEmbryo, Ringers2023NovelEpithelia}.
With $\delta=5$ \textmu m, our cilia stress corresponds to a maximum force density of
$f_c=507$ N/$\mathrm{m^3}$, which is very similar to
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$~\cite{Siyahhan2014FlowVentricles}.
Another study that used a force density to model ciliary motion is
Thouvenin~\emph{et al.}~\cite{Thouvenin2020OriginCanal}, where two-dimensional flow
in the central canal of zebrafish embryos was modeled using a cilia force density of
4000 N/$\mathrm{m^2}$. Averaging this value over the central canal diameter 8.9 \textmu m
they measured yields 449 N/$\mathrm{m^3}$. It is very interesting to note the similarity 
in cilia force densities estimated in our and these two previous studies.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b, c.} Particle tracking data visualizing CSF flow directionality (\textbf{b}) and
    velocity magnitudes (\textbf{c}) in the middle ventricle.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue). Dorsal (D), ventral (V), anterior (A), posterior (P),
    left (L) and right (R) coordinates indicated.
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-driven/no-cardiac flow model.
    \textbf{f.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries. Time instants of streamline plots indicated.
    \textbf{g.} Streamlines of the CSF flow field simulated with the cardiac-induced/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{h--j.} Streamlines simulated with the baseline flow model (cilia+cardiac)
    at a time 25\% (\textbf{h}), 50\% (\textbf{i}) and 75\% (\textbf{j})
    into the cardiac cycle. %t=0.112613 s %t=0.225225 s %t=0.315315 s
    Scale bars 50 \textmu m.
    }
    \label{fig:fig1}
\end{figure}

@@

Proteins, neurotransmitters and other molecular substances
intrinsically diffuse within the brain ventricles, but are also
advected by the flow of CSF. A natural question is which of these
mechanisms that dominate for any given type of solute. To address this
question, we simulate the distribution and evolution of three
different diffusing solutes injected within a region of interest (ROI 1) in the
dorsal region of the middle ventricle (\Crefpanels{fig:fig2}{a}),
subject to the CSF flow induced by motile cilia and cardiac pressure
pulsations. In particular, we consider three different molecular
diffusion coefficients, corresponding to a large extracellular vesicle (EV, $D_1 = 2.17
\times 10^{-12}$ m$^2$/s), a 90 kDa Starmaker-Green Fluorescent Protein
(STM-GFP, $D_2 = 5.75 \times 10^{-11}$ m$^2$/s),
and a 26 kDa photoconvertible protein
Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).

@@

For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have reduced
mean concentrations. To summarize, our results show that the
ventricular geometry and the cross-sectional area of the ventricular ducts matter
for solute transport, with notable differences in transport resulting from
relatively modest geometry modifications. This may explain why
we observed larger differences in the cilia-deficient computational
model results compared to the experimental data in \Cref{fig:fig4},
as the model and the fish have slightly different geometries.
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios display $F_i/F_{\mathrm{original}}\times 100\%$.}
    \label{fig:fig6}
\end{figure}

@@

\subsection{Numerical approximation and computational modeling aspects}
We modeled motile cilia as a stress
exerted tangentially on the CSF at the ventricular walls
while employing a slip boundary condition,
motivated by experimental observations of the
flow field in zebrafish embryo brain ventricles, both previously and in this work, in which
the fluid velocity increases towards the
wall~(\Crefpanels{fig:fig1}c)~\cite{Olstad2019CiliaryDevelopment}. 
The Brezzi-Douglas-Marini and discontinuous Galerkin scheme
$\mathrm{BDM}_1$-$\mathrm{DG}_0$
we used to discretize the governing equations of the flow
is computationally more expensive
than more conventional approaches such as a
Taylor-Hood discretization
($P_k-P_{k-1}$ elements)~\cite{Stenberg1990ErrorProblem}, or lower-order enriched
elements, such as the mini-element
($P_1-P_1$ with a velocity stabilization)~\cite{Brezzi2011MixedMethods}.
Here, $P_k$ is the space of continuous Lagrange polynomials of degree less
than or equal to the integer $k$.
We opted for the BDM-DG scheme because of its exactly-mass-conserving property;
by the definition of the elements, the numerical approximation of
the CSF velocity $\uu_h$ satisfies the continuity equation (\Cref{eq:stokes_eqs_cont}).
This yields $\normlinf{\nabla\cdot\uu_h}=0$ to within solver tolerances,
an important property for the stability of the advection-diffusion
equation\cite{cesmelioglu2022compatible, johnson2009numerical, gresho1998incompressible}.
In addition to its favorable mass-conserving property,
the BDM-DG discretization scheme 
we employed for the Stokes equations
lets us impose the impermeability condition $\uu\cdot\nn=0$
strongly without any tailored approach,
since the BDM elements control the normal component of the 
velocity~(\Cref{subsec:appendixA1})~\cite{Brezzi1985TwoProblems}.
If instead (for example) Taylor-Hood
elements were to be used, the impermeability boundary condition must
be handled with for example a Nitsche method~\cite{Nitsche1971UberSind}
or with a Lagrange multiplier~\cite{Babuska1973TheMultipliers,
Bertoluzza2017BoundaryHemodynamics}, which affects
convergence properties and increases problem complexity.

The ciliated boundaries in our CSF flow model are regionally homogeneous,
whereas in reality, cilia populate the ventricle walls in a
heterogeneous manner~\cite{Olstad2019CiliaryDevelopment}. Heterogeneous
arrangement of cilia, both in location, beating direction and beating frequency,
has been shown to enhance particle clearance in airway cilia arrays
in mice~\cite{Ramirez-SanJuan2020Multi-scaleArrays}.
Incorporating spatial arrangement heterogeneity in our model could
result in different transport timescales,
but we would not expect notable changes to the general trends of the macroscopic
transport that we study here, since the effects of introducing heterogeneous spacing
of cilia would alter the magnitude of cilia forces, but not their direction.
On the other hand, introducing more heterogeneous beating directionality could
alter the transport timescales as well as the directionality.
In our model, all of the applied cilia
tangential traction is aligned with the rostrocaudal axis, and more heterogeneity
in this alignment to reflect heterogeneity in the beating directionality of
cilia would presumably affect both magnitude and directionality of advective transport.
Previous studies have demonstrated that changes to beating directionality affects transport,
and that specific coordination of the ciliary beating is important in the brain ventricles for
flow and signaling processes~\cite{Eichele2020Cilia-drivenVentricle, Olstad2019CiliaryDevelopment,
Faubel2016Cilia-basedVentricles, Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Pellicciotta2020Cilia},
but also in other phenomena such as maintaining healthy functioning of mucous transport in
airways~\cite{Rayner1996CiliarySyndrome, Schneiter2021Multi-scaleFunction,
Bustamante-Marin2019LackClearance, Tsukita2012CoordinatedFeet} and in nodal flow which establishes
left-right symmetry in mammals~\cite{Yoshiba2014RolesSymmetry, Hirokawa2006NodalAsymmetry,
Sawamoto2006NewBrain}.
Although these studies were mostly on spatial scales resolving individual cilia,
we could envision the same effects in our continuum model,
similar to the results of previous numerical studies~\cite{Ramirez-SanJuan2020Multi-scaleArrays,
Thouvenin2020OriginCanal, Yoshida2022EffectVentricles}.

To simulate protein transport in the brain ventricles, we numerically solved
an advection-diffusion equation. Except for where we imposed a normal pressure
boundary condition on the CSF flow,
we imposed a no-flux condition on the concentration. The normal pressure
boundary condition leads to in- and outflow of CSF,
and this advective transport requires handling of in- and outflux boundary conditions on
the concentration. For the influx, we applied an approximate periodic boundary condition
that conserves the total solute concentration.
We handled the outflux by imposing zero diffusive flux while
leaving the advective flux term as an unknown in the weak form
of the advection-diffusion equation. This approach has been referred to with different names,
such as the ``free boundary condition"~\cite{Papanastasiou1992ACondition},
the ``no boundary condition"~\cite{Griffiths1997TheCondition}, and the 
``consistent flux boundary condition"~\cite{Lynch2020NumericalHemodynamics}. 
In leaving the advective flux term in the weak form we are lacking a boundary condition,
which is problematic in the continuous sense of the problem.
However, the discrete problem is well-posed, because the approach represents
constraining higher-order derivatives of the concentration~\cite{Griffiths1997TheCondition}. 
In defining ROIs 5 and 6 in the anterior and posterior ventricles,
we deliberately excluded the space proximal to
in- and outflow boundaries to limit adverse effects of 
the advective flux boundary condition on the mean concentrations.

A possible improvement to the in- and outflow boundary conditions would be
to consider a 3D-0D coupling of in- and outflow using
ordinary differential equations,
similar to applications in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
However, if the brain ventricular system in reality is a closed system,
one should rather seek out a way of modeling the CSF flow without open boundaries to avoid
handling in- and outflow boundary conditions. A more appropriate way of handling the beating
cardiac motion could be to impose a ventricular wall deformation, assumed to result from pulsatile
motion of the heart and/or cardiac vessels surrounding the ventricles. This wall deformation would
induce CSF flow while simultaneously permitting enforcement of the impermeability condition
$\uu\cdot\nn=0$ on the ventricular walls. Such wall motion-induced CSF flow has been modeled and
simulated successfully in the human brain ventricular system~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

We heuristically determined the magnitude of the cilia tangential traction boundary condition
with the parameter $\tau = 0.65$ mPa. The maximum stress
$\vert\btau\vert_{\mathrm{max}} = 2.5 \ \mathrm{mPa} = \num{2.5e-3} \ \mathrm{N/m^2}$
was applied in the dorsal-posterior region of the middle ventricle.
Previous simulation studies of cilia-mediated flow in brain ventricles have
used a volumetric force density to represent the cilia.
Based on reconstruction of CSF flow in murine brain ventricles,
a previous study employed a force that increased linearly from
the ventricular wall to a maximum force density of
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$ at a distance 15 \textmu m (the cilium length)
away from the wall~\cite{Siyahhan2014FlowVentricles}.
If we assumed a cilia array of the same thickness
$\delta = 15$ \textmu m in our model, the maximum stress $\vert\btau\vert_{\mathrm{max}}$
would correspond to a maximum force density of 
\begin{equation*}
    f_c = \frac{\vert\btau\vert_{\mathrm{max}}}{\delta}
        = \frac{\num{2.5e-3} \ \mathrm{N/m^2}}{\num{15e-6} \ \mathrm{m}}
        = 169 \ \mathrm{N/m^3},
\end{equation*}
which is smaller than $f_{\mathrm{max}}$, albeit the same order of magnitude.
The length 15 \textmu m is the typical length of cilia in the human brain
\cite{Afzelius2004CiliaRelatedDiseases, Siyahhan2014FlowVentricles},
while in zebrafish they are closer to 5 \textmu m~\cite{Olstad2019CiliaryDevelopment,
Salman2022ComputationalEmbryo, Ringers2023NovelEpithelia}.
With $\delta=5$ \textmu m, our cilia stress corresponds to a maximum force density of
$f_c=507$ N/$\mathrm{m^3}$, which is very similar to
$f_{\mathrm{max}} = 526$ N/$\mathrm{m^3}$~\cite{Siyahhan2014FlowVentricles}.
Another study that used a force density to model ciliary motion is
considered two-dimensional flow in the central canal of zebrafish
embryos~\cite{Thouvenin2020OriginCanal}.
Cilia was modeled using a cilia force density of
4000 N/$\mathrm{m^2}$. Averaging this value over the central canal diameter 8.9 \textmu m
measured in the study yields a maximum force of 449 N/$\mathrm{m^3}$.
It is very interesting to note the similarity 
in cilia force densities estimated in our and these two previous studies.
\fixme{Previous work concluded that motile cilia align by response to the shear
stress of flow. Perhaps there is an optimal cilia stress?}

The diffusion coefficients used in this study are based on experimental values. 
The value $D_3=\num{1.15e-10} \ \mathrm{m^2/s}$
is based on the value $\num{1.15e-10} \pm 0.11 \ \mathrm{m^2/s}$ reported in an
experiment~\cite{GuraSadovsky2017MeasurementExpansion}, while the Stokes-Einstein
relation~(\Cref{eq:D_stokes_einstein})~\cite{Einstein1905UberTeilchen} was used to extrapolate
the value of $D_2$ and estimate the value of $D_1$, based on experimental
data~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells,
Mahmood2023ExosomeTemperature}. The Stokes-Einstein relation assumes the diffusing
substance is of spherical shape, which proteins are not, and the relation does not
necessarily hold in CSF. Moreover, the extrapolation implies the assumption that
the two substances (the one for which we determine $D$ and the one we extrapolate from)
have equal densities. Consequently, there is uncertainty in the diffusion coefficient values used.
We note that one previous study reported that
Dendra2 satisfies the Stokes-Einstein in a solution of similar viscosity
to CSF\cite{GuraSadovsky2017MeasurementExpansion}.
We used CSF density and viscosity values measured in
humans~\cite{Bloomfield1998EffectsFluid},
and because embryonic CSF in zebrafish is high in proteins, nutrients and other substances,
the CSF in which the photoconverted Dendra2 are transported in our experiments may be more viscous
than we assumed. This could reduce the diffusion coefficient (cf.~\Cref{eq:D_stokes_einstein}),
but based on reports of protein concentration having a limited impact on the viscosity of
CSF~\cite{Bloomfield1998EffectsFluid, Brydon1995PhysicalViscosity}, this reduction is presumably
limited. Merely the presence of other proteins and nutrients may also reduce the diffusion
coefficient by reduction of mean-free path lengths of molecules~\cite{Cussler2009Diffusion:Systems}.
Considering the diffusion coefficients $D_1$ and $D_2$ (both smaller
than $D_3$) in our transport analysis provides information about how the simulation
results of Dendra2 transport would be altered if the
diffusion coefficient $D_3$ is overestimated. In the case that $D_3$ is underestimated, 
we expect a trend similar to the difference in the $D_2$ and $D_3$ solute concentrations;
incerasingly diffusive fields.
